---
pagetitle: 'A brief guide to the **`dbicc`** package'
author: 'Meng Xu'
output:
  html_document:
    df_print: paged
date: "2020/6/4"
vignette: >
  %\VignetteIndexEntry{README}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# A brief guide to the **`dbicc`** package

This package implements a new approach to reliability, extending the classical intraclass correlation coefficient (ICC) to a new measure based on arbitrary distances among observations (Xu et al., 2020).

The key function `dm2icc` computes the reliability (distance-based ICC, or dbICC) for a given matrix of distances among observations, while `dm2icc.bt` gives boostrap confidence intervals for the dbICC value.

The dataset `dt2v3` is the fMRI data set analyzed by Xu et al. (2020). `plotdmat` visualizes the associated distance matrix.

Please feel free to contact Meng Xu <mxu@campus.haifa.ac.il> with questions and bug reports.

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Installation

This package can be installed from GitHub, and loaded, as follows:
```{r install}
# devtools::install_github("wtagr/dbicc", force = TRUE)
library(dbicc)
```

## Distance matrix computation

The package does not include functions to compute the distance matrix, but the following function performs this step for our application. It inputs a *list* object and outputs a distance matrix determined by the *method* argument: `method = 'f'` yields the Frobenius norm, `method = 'r'` the $\sqrt{1-r}$ distance, and `method = 'l1'` the $\ell_1$ distance (sum of absolute differences).

```{r mdist}
mdist <- function(datalist, method=c("f")) {
    dmat<-matrix(0,length(datalist),length(datalist))
    for (i in 2:length(datalist)) for (j in 1:(i-1)){
        if (method=="r"){
            v1<-datalist[[i]][lower.tri(datalist[[i]], diag = FALSE)]
            v2<-datalist[[j]][lower.tri(datalist[[j]], diag = FALSE)]
            r<-cor(v1,v2)
            dmat[i,j]=dmat[j,i]=sqrt(1-r)
        } else if (method=='l1'){
            dmat[i,j]=dmat[j,i]=sum(abs(datalist[[i]]-datalist[[j]]))
        } else {
            dmat[i,j]=dmat[j,i]=norm(datalist[[i]]-datalist[[j]],type=method)
        }
    }
    return(dmat)
}
```


`dt2v3` contains fMRI time series for 25 individuals with 2 scans each, grouped by individual. Here we compute a matrix of distances among the correlation matrices resulting from these multivariate time series:

```{r,  fig.height = 5, fig.width = 7, fig.align = "center"}
data(dt2v3)

# compute correlation matrices
cor2v3 <- lapply(dt2v3,cor)

# compute the matrix of distances among the correlation matrices
distmat <- mdist(cor2v3)

# visualize
plotdmat(distmat, 25, 2)
```



## Compute the reliability  

We can then input the distance matrix to `dm2icc`, to compute the dbICC:
```{r dbicc}
dm2icc(distmat,25,2)
```

Obtain confidence interval for the dbICC, based on 100 boostrap replicates (a larger number of replicates is recommended in practice):
```{r Bootstrap CI}
dm2icc.bt(distmat,25,2, nB=100)
```

## Reference
Xu, M., Reiss, P. T., and Cribben, I. (2020). Generalized reliability based on distances. *Biometrics*, to appear.
