---
pagetitle: 'A brief guide for the package **`dbICC`**'
author: 'Meng Xu'
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Application-vector_type}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
date: "2020/3/15"
---

# A brief guide for the package **`dbICC`**

This package implements a new approach to measuring the *test-retest* reliability by extending the classical intraclass correlation coefficient (ICC) via arbitrary distances among observations. See Xu et al. (2019) for details.

The key function `dm2icc` gives a value of reliability by inputing size of sample and related *distance matrix*. And `dm2icc.bt` gives (un)corrected Boostrap Confidence Interval of dbICC value.

The dataset `dt2v3` is a sample of fMRI data set, permformed in the paper Xu et al. (2019). `plotdmat` visualizes the related distance matrix.

Please feel free to contact Meng Xu <mxu@campus.haifa.ac.il>, if you have some questions and find bugs.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Installation

This package is available only on GitHub now. You can intall it through this:
```{r install}
# devtools::install_github("wtagr/dbicc", force = TRUE)
library(dbicc)
```

## Preliminary: Distance Matrix

In the package, we do not contain any function on acquirng the distance matrix. But here is a sample distance function based on different distances and norms via the input of a *list* object.

```{r mdist}
mdist <- function(datalist, method=c("f")){
  dmat<-matrix(0,length(datalist),length(datalist))
  for (i in 2:length(datalist)) for (j in 1:(i-1)){
    if(method=="r"){
      v1<-datalist[[i]][lower.tri(datalist[[i]], diag = FALSE)]
      v2<-datalist[[j]][lower.tri(datalist[[j]], diag = FALSE)]
      r<-cor(v1,v2)
      dmat[i,j]=dmat[j,i]=sqrt(1-r)
    }else if(method=='sl1'){
      dmat[i,j]=dmat[j,i]=sum(abs(datalist[[i]]-datalist[[j]]))
    }else{
      dmat[i,j]=dmat[j,i]=norm(datalist[[i]]-datalist[[j]],type=method)
    }
  }
  return(dmat)
}
```
Note that `method = 'f'` denotes the Frobenius norm, `method = 'r'` denotes the $\sqrt{1-r}$ distance, and `method = 'sl1'` denotes the $\ell_1$ distance (sum of absolute differences).


Set `dt2v3` as an example:

```{r,  fig.height = 5, fig.width = 5, fig.align = "center"}
load("dt2v3.rda") # data(dt2v3)
# calculate correlation matrices
cor2v3 <- lapply(dt2v3,cor)

# calculate the distance matrix
distmat <- mdist(cor2v3)

```

Note that `dt2v3` has 25 subjects with 2 scannings each. And the order of the input data list is grouped by the subject.

## Calculate the reliability

Plug in the *distance matrix* to `dm2icc`:

```{r dbicc}
dm2icc(distmat,25,2)
```

Get its Confidence Intervel based on 100 boostraps:
```{r Bootstrap CI}
dm2icc.bt(distmat,25,2, nB=100)
```

